On Adding Pattern Matching to Haskell-based Deeply Embedded Domain Specific Languages

### 2020-12-02 paper/paper.tex
Type equality syntactic clarification
Some description of papers cited in motivation and some reformatting to fit page limit

\newcommand{\typeeq}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}\,\;}

@@

A major benefit of the Haskell language is that it is frequently possible to derive
a Haskell implementation from a specification, either partially or fully.  This
technique is used extensively, for example, in \cite{Bird:2010:Pearls} and
\cite{Gibbons:2002:Calc}.  Recent examples of this technique include the work in
\cite{Elliott-2018-ad-icfp} as well as the work in
\cite{Elliott2019-convolution-extended}.  When it is not possible to fully
derive an implementation from a specification, it is often easier to check a
Haskell implementation against a specification than it is in many imperative
languages due, in part, to the management of side effects in Haskell. This is
more difficult to do using an imperative language such as C, as they provide no
restrictions on where side effects can occur.

@@

\noindent The equality constraint ensures that a canonical type is its own canonical type.

@@

namesEq :: forall a b.  (Typeable a, Typeable b) =>
  Name a -> Name b -> Maybe (a :~: b)
namesEq (Name n) (Name n') =
  case eqT :: Maybe (a :~: b) of
    Just Refl
      | n == n'   -> Just Refl
      | otherwise -> Nothing
    Nothing       -> Nothing
\end{lstlisting}

@@

The plugin's control flow is described in Figure~\ref{fig:CorePlugin}. When control
flow reaches a node, the rewrite rules listed in parentheses are performed. These rewrite
rule names correspond to the rules given in Figure~\ref{fig:Rewrites}.
%
In Figure~\ref{fig:Rewrites},
$\varepsilon$ represents an empty term (in the context of the term rewriting system used in that Figure). Calls
to \ttt{unrep :: E a -> a} are only used internally and will not exist in
the final result of the transformation.

\begin{figure}
  \centering
  \resizebox{0.8\columnwidth}{8.2cm}{\includegraphics{Diagrams/Flowchart.pdf}}
   \caption{Control Flow of the Core Plugin}
   \label{fig:CorePlugin}
\end{figure}%

\pagebreak
Note that:
\begin{itemize}
  \item The tail recursion transformation given by $D$, $T'_{\ttt{f}}$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type. As a result, the tail recursion transformation can be used on its own.

>>>>

\newcommand{\typeeq}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}\,\;}
\newcommand{\typeeqb}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}}

@@

A major benefit of the Haskell language is that it is frequently possible to
derive a Haskell implementation from a specification, either partially or fully.
In \cite{Bird:2010:Pearls} a collection of algorithms is presented with implementations derived in this
manner. The work \cite{Gibbons:2002:Calc} describes many of the
mathematical tools underlying such program derivations.
Recent examples of this technique include the work in
\cite{Elliott-2018-ad-icfp}, where an implementation of automatic
differentiation is derived from a specification
of derivatives.  Similarly, the work in
\cite{Elliott2019-convolution-extended} derives an implementation of a
generalized form of convolution from a specification.  When it is not possible to fully derive an implementation, it is often easier to check a Haskell implementation against a
specification than it is in many imperative languages due, in part, to the
management of side effects in Haskell. This is more difficult to do using an
imperative language such as C, as they provide no restrictions on where side
effects can occur.

@@

\noindent The equality constraint \ttt{(\typeeqb)} ensures that a canonical type is its own canonical type.

@@

namesEq :: forall a b.  (Typeable a, Typeable b) =>
  Name a -> Name b -> Maybe (a :~: b)
namesEq (Name n) (Name n') =
  case eqT :: Maybe (a :~: b) of
    Just Refl
      | n == n' -> Just Refl
    _           -> Nothing
\end{lstlisting}

@@

The plugin's control flow is described in Figure~\ref{fig:CorePlugin}. When control
flow reaches a node, the rewrite rules listed in parentheses are performed. These rewrite
rule names correspond to the rules given in Figure~\ref{fig:Rewrites}.
%
In Figure~\ref{fig:Rewrites},
$\varepsilon$ represents an empty term (in the context of the term rewriting system used in that Figure). Calls
to \ttt{unrep :: E a -> a} are only used internally and will not exist in
the final result of the transformation.
%
\begin{figure}
  \centering
  \resizebox{0.8\columnwidth}{8.2cm}{\includegraphics{Diagrams/Flowchart.pdf}}
   \caption{Control Flow of the Core Plugin}
   \label{fig:CorePlugin}
\end{figure}%

% \pagebreak
Note that:
\begin{itemize}
  \item The tail recursion transformation given by $D$, $T'_{\ttt{f}}$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type. As a result, the tail recursion transformation can be used on its own.
